Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-02-26 paper.tex
update

Note that the store is monotonic. TODO

@@

The co-domain of binding stores \texttt{Storable} is the same
as we defined for CESK machines.
The co-domain of continuation stores \texttt{Cont} is comprised of 
a frame \texttt{Frame} and a continuation address \texttt{KAddr}.
To mimic the runtime call stack, \texttt{KAddr} plays the role of
representing the remaining stack frames.
TODO: nondet of cont?

@@

atomicEval returns a set of storable values, i.e., closures.

\begin{verbatim}

def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
  case Var(x) => bstore(env(x))
  case lam@Lam(x, body) => Set(Clos(lam, env))
}
\end{verbatim}

@@

However, comparing with concrete CESK machine, the
\texttt{drive} function performs collecting semantics instead of valuation
semantics. That is, for the purpose of analyzing programs, the function \texttt{drive}
collects all the intermediate machine states as the program is abstractly executing.
The following code shows a variant of work list algorithm that always applies \texttt{step}
function to the first element \texttt{hd} of the work list \texttt{todo} if \texttt{hd}
is unseen, then adds the result of \texttt{step} to the work list and also adds \texttt{hd}
to the explored states set.
If the work list is empty, the function just
returns the set of explored states up to now.

@@

In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
the return-flow of invocation \texttt{(id 2)} 
is going to both call-sites \texttt{(id 1)} and \texttt{(id 2)}. Therefore, the
returned value \texttt{2} for variable \texttt{y} is also propagated to 
variable \texttt{x} and then imprecise analysis result is arose.
This return-flow merging is inevitable even we increase the context-sensitivity.
If we use the monovariant analysis 0CFA, the analyzing result would be \texttt{x} and 
\texttt{y} point to set \texttt{\{1, 2\}}, because it not distinguished 
\texttt{z} at different call-sites.
Under $1$-CFA, the algorithm is able to distinguish that variable \texttt{z} of function
\texttt{id} has two different values at two call-sites, so variable \texttt{y} 
would not be polluted by \texttt{1}, but it is still helplessness for distinguish
return-flows.
TODO: do need more details to explain this example? (as in P4F paper)

In this section, we firstly describe a simple but uncomputable solution to the
problem; then briefly review several recent research works that enable
computable and precise call/return match. Even though the first solution is
uncomputable, but as we will see in section 3, it is the starting point of our
transformation.

@@

the only non-determinism in our state transition function is when dereferencing
the callee \texttt{f} from store at a call-site.

@@

On the other side, with the insight from Reynold that the defined language will inherit properties
from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
definitional interpreters (ADI) that automatically inherit the pushdown control-flow 
property from the defining language.
TODO: connection to this paper.

Question: although they establish pushdown control-flow analysis in different ways, 
is there a correspondence between different methods?

In the rest of this section, we take a closer look at the P4F and ADI which are
two state-of-art techniques to accomplish perfect call-stack precision.

@@

@@

>>>>

TODO(rephrase): Note that the store is monotonic, the store is always growing and never becoming 
smaller. This property guarantees we eventually will find a fixed point.

@@

The co-domain of binding stores \texttt{Storable} is the same
as we defined for CESK machines.
The co-domain of continuation stores \texttt{Cont} is comprised of 
a frame \texttt{Frame} and a continuation address \texttt{KAddr}.
To mimic the runtime call stack, \texttt{KAddr} plays the role of
representing the remaining stack frames.
TODO: talk about nondet of cont deref?

@@

Since dereferencing an address becomes non-deterministic, our \texttt{atomicEval}
function also truns into non-deterministic. Given an atomic expression \texttt{e},
\texttt{atomicEval} returns a set of storable values, i.e., closures, to the caller.
If the expression is just a \texttt{lambda} term, then the returned set is singleton.

\begin{verbatim}
def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
  case Var(x) => bstore(env(x))
  case lam@Lam(x, body) => Set(Clos(lam, env))
}
\end{verbatim}

The structure of function \texttt{step} is similar with concrete CESK machines, 
except non-determinism which makes \texttt{step} returns a list of possible
successor states.
In the beginning of \texttt{step}, we call function \texttt{tick} to update 
the time.
Then we have two cases to consider: 1) Current control expression is \texttt{Let},
and \texttt{App(f, ae)} is being bind to \texttt{x}.
Fristly we retrieve the set of closures that \texttt{f} may represents.
For each closure in the set, we do almost the same as what we did in concrete 
CESK machines. But note that since the continuation is allocated on store (\texttt{kstore}),
a new continuation address \texttt{new_kaddr} is constructed, and a new frame
\texttt{Frame(x, e, env)} paired with the current continuation address \texttt{kaddr}
is joined into this address. Finally, a list of states is generated.

2) An atomic expression \texttt{ae} sits on the control position of the state.
The value of \texttt{ae} is being returned to its caller. 
We dereference the continuation address \texttt{kaddr} and obtain a set of
continuations \texttt{conts}.
For each continuation in the set, we construct an environment based on the
environment of the frame (\texttt{env_f}), and bind \texttt{x} to a newly created binding 
address \texttt{baddr}; also update the store with \texttt{baddr} and 
the values that \texttt{ae} represents. In every generated states,
the control becomes the expression \texttt{e} in the frame,
and as we can tell from the name, continuation address \texttt{f_kaddr} 
also comes from the frame.

@@

However, comparing with concrete CESK machine, the
\texttt{drive} function performs collecting semantics instead of valuation
semantics. That is, for the purpose of analyzing programs, the function \texttt{drive}
collects all the intermediate machine states as the program is abstractly executing.
The following code shows a variant of work list algorithm that always applies \texttt{step}
function to the first element \texttt{hd} of the work list \texttt{todo} if \texttt{hd}
is unseen, then adds the result of \texttt{step} to the rest of work list and in 
the meanwhile adds \texttt{hd} to the explored states set.
If the work list is empty, the function just returns the set of reachable states up to now.

@@

In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
the return-flow of invocation \texttt{(id 2)} 
is going to both call-sites \texttt{(id 1)} and \texttt{(id 2)}. Therefore, the
returned value \texttt{2} for variable \texttt{y} is also propagated to 
variable \texttt{x} and then imprecise analysis result is arose.
This return-flow merging is inevitable even we increase the context-sensitivity.
If we use the monovariant analysis 0CFA, the analyzing result would be \texttt{x} and 
\texttt{y} point to set \texttt{\{1, 2\}}, because it not distinguishes
\texttt{z} at different call-sites.
Under $1$-CFA, the algorithm is able to distinguish that variable \texttt{z} of function
\texttt{id} has two different values at two call-sites, so variable \texttt{y} 
would not be polluted by \texttt{1}, but it is still helplessness for distinguish
return-flows.
TODO: do need more details to explain this example? (as in P4F paper)

In this section, we firstly show a simple but uncomputable solution to the
problem; then briefly review several recent research works that enable
computable and precise call/return match. Even though the first solution is
uncomputable, but as we will see in section 3, it is the starting point of our
transformation.

@@

The state transition function \texttt{step} is still non-determinism, but
the only non-determinism happens when dereferencing the callee \texttt{f} 
from the store at call-site \texttt{App(f, ae)}. 
For each closure of \texttt{f} in the set, we do exactly
the same as what we did in concrete CESK machines: a new frame is constructed
and pushed onto the stack.
In the second case, since the continuation \texttt{konts} is a list of 
frames, we can simply decompose it to differentiate an empty stack
or a frame exists on the top of stack. 
If the stack is empty, then empty successors is returned, because the computation
is finished.
Otherwise, we matched a frame \texttt{Frame(x, e, env_f)}. 
This transition is deterministic, because we only have one frame, this frame
is then popped and the new state uses the rest of stack.

@@

In the rest of this section, we take a closer look at the P4F which is the
state-of-art techniques to accomplish perfect call-stack precision on small-step
AAMs, and Abstracting De nitional Interpreters (ADI) that adpots another 
approach to pushdown control flow analysis.

@@

On the other side, with the insight from Reynold that the defined language will inherit properties
from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
definitional interpreters (ADI) that automatically inherit the pushdown control-flow 
property from the defining language.
TODO: connection to this paper.
TODO: although they establish pushdown control-flow analysis in different ways, 
is there a correspondence between different methods?

@@

\subsubsection{First Try}

simply trun evaluation contexts into higher-order continuations

\subsubsection{Computable Refunctionalization}

fix point cache
